### 2021-05-18 manuscript/introduction.tex

\section{Introduction}

Equivalent layer background:

\cite{dampney1969}, \cite{blakely1996}, \cite{emilia1973}, \cite{li2014using}
\\\\
Fast methods for equivalent layer:

\cite{leao-silva1989}, \cite{mendonca-silva1994}, \cite{oliveirajr-etal2013}, \cite{li-oldenburg2010}, \cite{siqueira-etal2017}, \cite{mendoncca2020subspace}
\\\\
Methods for solving Toeplitz matrices:

\cite{golub-vanloan2013}, \cite{levinson1946}, \cite{chan-jin2007}
\\\\
Toeplitz matrices in Potential methods:

\cite{zhang-wong2015}, \citealp{zhang-etal2016}, \cite{hogue2020tutorial}, \cite{renaut2020fast}
\\\\
Convolutional equivalent layer grav:

\cite{siqueira-etal2017}, \cite{takahashi2020convolutional}
\\\\
About this work: